### 2022-08-24 article.tex
Progress

To \emph{obtain} the raw genetic data, there may have been fieldwork and/or lab-work
involved. For example, a researcher needs to go into the field to take
samples (e.g. blood) of humans or other species. 
To analyse these samples,
a researcher needs a lab. 
Depending on the analysis, there may be a big bioinformatics step
to aggregate the measurements into usable data. 
Although this paper does not focus on the code run to 
aggragate the raw data into useful genetic data, the same arguments
can be made for that code as in the example: that code is 
the ground truth of how the raw data is collected into useful genetic data.

@@

Programming code (from now on: 'code') is paradata, 
as it is data that describes how data is generated.
Code is data that is usually in the form of text, 
spread over one or more files, that describes the experiment.
The computational experiment (from now on: 'experiment') collects the data
and then looks for patterns.
Those patterns (again, a type of data) are the results of an experiment.

The (scientific) article is metadata, as it is data that describes other data:
an article describes the
experiment (and hence the programming
code) in a natural language, which is usually English (at this point in time). 
However, a natural language is not the best
candidate to describe how the data is collected,
as it has a loose connection with collecting the data.
If the code and article of an experiment disagree,
it is the code that actually let the data be collected.  
Instead, an article is metadata about a research.

@@

A scientific claim must be reproducible before it is accepted
by the scientific community.
Reproducibility (i.e. to reproduce the same results) 
and replicability (i.e. to reconclude a conclusion)
are fundamental characteristics of scientific studies \cite{patil2019visual}.
For computational science, it is relatively easy to 
reproduce an experiment, as all it takes is a computer, electricity,
an optional internet connection, the code and the data.
In computational science, however, it appears that 
culture to reproduce results has been lost 
and to counter this, it has been suggested to make
reproducible research a minimal requirement for 
publication \cite{peng2011reproducible}.

@@

The code of a computational experiment must be published
for an experiment to reproducible.
It is however (too) common that the original data
and/or code are unavailable \cite{peng2021reproducible}.
One saddening example is described in \cite{haibe2020importance}, where
an algorithm that detects breast cancer better than a human expert,
was irreproducible.
In the case of genetic epidemiology, it is a given that the original
data cannot be published as this data is sensitive.
The code of a computational experiment, however, \emph{can} be
distributed without such problems.
It is, however, not code that is published, but an English
description of what that code does instead.
From a knowledge management's perspective,
to conserve an experiment,
it is the code that must be archived and not the
English description of what it does:
it is the code does the actual work.
Hence, from a knowledge management perspective, emphasis should be
put more on the preservation of code, as it is the most important actor
in an experiment.

@@

Code should be published, as it holds the
ground truth of an experiment; it does the actual work.
The more complex the computation pipeline is, the easier it is
to have a mismatch between the article (that describes what the
code does) and the code (that actually does the work).
It is easy (tempting?) to overlook how easy it is to have a mismatch
between paper and code.

To illustrate how easy it is to get a mismatch between a paper
and the code, consider this fictional example
of a text in a paper:

>>>>

To \emph{obtain} the raw genetic data, there may have been fieldwork and/or lab-work
involved. For example, a researcher needs to go into the field to take
samples (e.g. blood) of humans or other species. 
To analyse these samples,
a researcher needs a lab. 
Depending on the analysis, there may be a big bioinformatics step
to aggregate the measurements into usable data. 
Although this paper does not focus on the code run to 
aggragate raw and noisy measurements into usable data, the same arguments
can be made for that code as in the example: that code is 
the ground truth of how the raw data is converted into useful genetic data.

@@

Programming code (from now on: 'code') is paradata, 
as it is data that describes how data is generated.
Code is data that is usually in the form of text, 
spread over one or more files, that describes an experiment/annalysis.
The computational experiment (from now on: 'experiment') collects the data
and then looks for patterns.
Those patterns (again, a type of data) are the results of an experiment.

The (scientific) article is metadata, as it is data that describes other data:
an article describes the
experiment; an article describes what the code does in a natural 
language, such as English (which I'll use from now on).
However, English (or any other natural language) is not the best
candidate to describe how the data is collected,
as it only has a loose connection with collecting the data.
If the code and article of an experiment disagree,
it is the code that actually let the data be collected.  
Instead, an article is metadata about a research.

@@

A scientific claim must be reproducible before it is accepted
by the scientific community.
Reproducibility (i.e. to reproduce the same results) 
and replicability (i.e. to reconclude a conclusion)
are fundamental characteristics of scientific studies \cite{patil2019visual}.
For computational science, theoretically, it should be relatively easy to 
reproduce an experiment, as all it takes is a computer, electricity,
an optional internet connection, the code and the data.
In practice, however, it appears that 
the academic culture to reproduce results 
has been lost over time \cite{peng2011reproducible}.
That same study recommended, as a solution to counter this, 
to make reproducible research a minimal requirement for 
publication \cite{peng2011reproducible}.

@@

The code of a computational experiment must be published
for an experiment to reproducible.
It is however (too) common that the original data
and/or code are unavailable \cite{peng2021reproducible}.
One saddening example is described in \cite{haibe2020importance}, where
an algorithm that detects breast cancer from images better than a human expert,
was irreproducible.
In the case of genetic epidemiology, 
the original data usually cannot be published, 
due to the guaranteed privacy of the people having shared their characteristics.
The code of a computational experiment, however, \emph{can} be
distributed without such problems.
In academic journals, however, its is not code that is published, 
but an English description of what it does instead.

From a knowledge management's perspective,
it is the code that must be archived 
to conserve an experiment,
and not the
English description of what it does:
it is the code does the actual work.
Hence, from that perspective, emphasis should be
put more on the preservation of code, as it is the most important actor
in an experiment.

@@

From a broader perspective, code should be published, as it holds the
ground truth of an experiment; it does the actual work.
The more complex the computation pipeline is, the easier it is
to have a mismatch between the article (that describes what the
code does) and the code (that actually does the work).
It is easy (tempting?) to overlook how easy it is to have a mismatch
between paper and code.

To illustrate how easy it is to get a mismatch between a paper
and the code, consider this example
of a text in a fictional paper:
